---
title: "House Price Compition"
author: "Ankita Gairola"
date: "2023-08-06"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### 1. Introduction

Goal: Estimate the sale price of property

Dataset: Data from Kaggle. In this dataset, 79 explanatory variables describing (almost) every aspect of residential homes in Ames, Iowa. 


#### 2. Data Assessment

##### 2.1 Set the Directory 

getwd()

setwd("E:/Kaggle/Competition_1)

##### 2.2 Install packages

install.packages("readr")

install.packages("ggplot2")

install.packages("gplots")

install.packages("repr")

install.packages("tidyverse")

install.packages("skimr")

install.packages("dplyr")

install.packages("plyr")

```{r}
library(readr)

library(ggplot2)

library(repr)

library(tidyverse)

library(gplots)

library(skimr)

library(dplyr)

library(plyr)
```


##### 2.2 Import Data

```{r}
Test <- read.csv("E:/Kaggle/Competition_1/house-prices-advanced-regression-techniques/test.csv")

Train <- read.csv("E:/Kaggle/Competition_1/house-prices-advanced-regression-techniques/train.csv")

```

##### 2.3 Exploring the Data

```{r}
colnames(Test)

colnames(Train)
```


```{r}
dim(Test)

dim(Train)
```

```{r}
skim_without_charts(Test)

skim_without_charts(Train)
```


```{r}
str(Test)

str(Train)
```

###### 2.3.1 Any missing value

###### Missing value in Rows

```{r}
missing_row_train <- Train[!complete.cases(Train),]

head(missing_row_train)

missing_row_test <- Test[!complete.cases(Test),]

head(missing_row_test)
```

```{r}
nrow(Train)

nrow(Test)

nrow(missing_row_test)

nrow(missing_row_train)
```

###### Combined Dataset

```{r}
Complete_Data <- merge(Test, Train, all = TRUE)

view(Complete_Data)

head(Complete_Data)

dim(Complete_Data)

```
###### Selecting important variable for working


```{r}
variable_name <- names(Complete_Data)

variable_name
```

Variables for new dataset

```{r}
 select_var <- c("Id", "MSZoning", "LotArea", "Utilities", "BldgType", "HouseStyle", "OverallQual", "OverallCond", "YearBuilt", "ExterQual", "ExterCond", "BsmtQual", "BsmtCond", "Heating", "HeatingQC", "CentralAir", "Electrical", "GrLivArea", "BedroomAbvGr", "KitchenAbvGr","KitchenQual",  "TotRmsAbvGrd", "Functional", "Fireplaces", "GarageType", "PoolArea","Fence","MoSold",   "YrSold", "SaleType", "SaleCondition", "SalePrice")
```


##### 2.4 New dataset for prediction

```{r}
new_dataset <- Complete_Data[, select_var]

head(new_dataset)
```

```{r}
summary(new_dataset)
```



```{r}
cleaned_data <- new_dataset[complete.cases(new_dataset), ]
```


###### 2.4.1 Analyze form sales price


```{r}
summary(cleaned_data$SalePrice)
```

###### Distribustion of sales price

```{r}
options(scipen=10000)
ggplot(cleaned_data, aes(x = SalePrice, fill = ..count..)) +
  geom_histogram(binwidth = 5000) +
  ggtitle("Figure 1 Histogram of SalePrice") +
  ylab("Count of houses") +
  xlab("Housing Price") + 
  theme(plot.title = element_text(hjust = 0.8))
```


###### Distribution of Sales price by MSZoing

```{r}
options(repr.plot.width=5, repr.plot.height=4)
ggplot(cleaned_data, aes(x = MSZoning, fill = MSZoning )) + 
geom_bar()+ 
scale_fill_hue(c = 80)+
ggtitle("Figure 2 Distribution of MSZoning")+
theme(plot.title = element_text(hjust = 0.5),legend.position="right", legend.background = element_rect(fill="grey90",
                                                                                                           size=0.5, linetype="solid", 
                                                                                                           colour ="black"))+
geom_text(stat='count',aes(label=..count..),vjust=-0.25)
```



###### Distribution of sale price by BldfType

```{r}
ddply(cleaned_data, .(BldgType), summarize,Total = length(BldgType),Max_price=max(SalePrice),Min_price=min(SalePrice))
```

```{r}
ggplot(cleaned_data, aes(SalePrice)) +
 geom_histogram(aes(fill = BldgType), position = position_stack(reverse = TRUE), binwidth = 20000) +
 coord_flip() + ggtitle("Figure 3 Histogram of SalePrice") +
 ylab("Count") +
 xlab("Housing Price") + 
 theme(plot.title = element_text(hjust = 0.5),legend.position=c(0.7,0.6), legend.background = element_rect(fill="grey90",
                                                                                                           size=0.6, linetype="solid", 
                                                                                                           colour ="black"))
```

###### Distribution of price by OverallQual

```{r}
ggplot(cleaned_data, aes(x = SalePrice,fill = as.factor(OverallQual))) +
  geom_histogram(position = "stack", binwidth = 10000) +
  ggtitle("Figure 4 Histogram of SalePrice") +
  ylab("Count") +
  xlab("Housing Price") + 
  scale_fill_discrete(name="OverallQual")+
  theme(plot.title = element_text(hjust = 0.7), legend.position=c(0.9,0.7), legend.background = element_rect(fill="grey90",
                                                                                                           size=0.7, linetype="solid", 
                                                                                                           colour ="black"))
```


##### 2.5. Corelation Exploreation

Correlation  between Sales price and TotRmsAbvGrd

```{r}
ggplot(cleaned_data, aes(x=TotRmsAbvGrd, y=SalePrice)) + 
  geom_point(shape=1) +  
  geom_smooth(method=lm , color="blue", se=FALSE)+
  ggtitle("Figure 5 Scatter plot of Sale Price and TotRmsAbvGrd") +
  theme(plot.title = element_text(hjust = 0.6))
```


Correlation  between Sales price and heating

```{r}
ggplot(cleaned_data, aes(x=Heating, y=SalePrice)) + 
  geom_point(shape=1) +  
  geom_smooth(method=lm , color="blue", se=FALSE)+
  ggtitle("Figure 6 Scatter plot of Sale Price and Heating") +
  theme(plot.title = element_text(hjust = 0.6))
```


Correlation  between Sales price and Lot Area


```{r}
ggplot(cleaned_data, aes(x=LotArea, y=SalePrice)) + 
  geom_point(shape=1) +  
  geom_smooth(method=lm , color="blue", se=FALSE)+
  ggtitle("Figure 7 Scatter plot of Sale Price and Lot Area") +
  theme(plot.title = element_text(hjust = 0.6))
```

Correlation  between Sales price and House Style

```{r}
ggplot(cleaned_data, aes(x=HouseStyle, y=SalePrice)) + 
  geom_point(shape=1) +  
  geom_smooth(method=lm , color="blue", se=FALSE)+
  ggtitle("Figure 8 Scatter plot of Sale Price and House Style") +
  theme(plot.title = element_text(hjust = 0.6))
```

Correlation between Sales price and OverallQual

```{r}
ggplot(cleaned_data, aes(x=OverallQual, y=SalePrice)) + 
  geom_point(shape=1) +  
  geom_smooth(method=lm , color="blue", se=FALSE)+
  ggtitle("Figure 9 Scatter plot of Sale Price and OverallQual") +
  theme(plot.title = element_text(hjust = 0.6))
```

Correlation between Sales price and GrLivArea

```{r}
ggplot(cleaned_data, aes(x=GrLivArea, y=SalePrice)) + 
  geom_point(shape=1) +  
  geom_smooth(method=lm , color="blue", se=FALSE)+
  ggtitle("Figure 10 Scatter plot of Sale Price and GrLivArea") +
  theme(plot.title = element_text(hjust = 0.6))
```

#### 3. Linear Regression Model

###### 3.1 Varibles for regression model

```{r}
model_variables <- c("Id", "OverallQual", "OverallCond", "YearBuilt", "ExterQual", "ExterCond", "BsmtQual", "BsmtCond", "HeatingQC", "CentralAir", "Electrical", "GrLivArea", "BedroomAbvGr", "KitchenAbvGr","KitchenQual", "TotRmsAbvGrd", "Functional", "Fireplaces", "GarageType", "PoolArea","Fence","SalePrice")
```


```{r}
Newmodel_Dataset <- new_dataset[, model_variables]

Newmodel_Dataset$SalePrice1 <- log(Newmodel_Dataset$SalePrice)
```

Dividing dataset in two parts. one is model_data_1 and other is model_data_2

```{r}
set.seed(10000)
model.index <- sample(c(1:dim(Newmodel_Dataset)[1]), dim(Newmodel_Dataset)[1]*0.6)
model_data_1 = Newmodel_Dataset[model.index,]
model_data_2 <- Newmodel_Dataset[-model.index,]


```

Run Regression

```{r}
Regdata <- lm(SalePrice1~.-SalePrice, data = model_data_1)
summary(Regdata)
```


###### 3.2 Classification of Tree

install.package("rpart")

install.packages("rpart.plot")

```{r}
library(rpart)

library(rpart.plot)
```



```{r}
Newmodel.tree <- rpart(SalePrice~.-SalePrice,
                    data = model_data_1,control = rpart.control(cp = 0.03))

plotcp(Newmodel.tree)
printcp(Newmodel.tree)
```



###### 3.3 Random Forest

install.packages("randomForest")

install.packages("yardstick")

```{r}
library(randomForest)

library(yardstick)
```
```{r}
model_data_3 <-na.omit(model_data_1)
```


```{r}
RF <- randomForest(SalePrice1 ~ . - SalePrice, data = model_data_3,
                   importance = TRUE, ntree = 500, nodesize = 7)
```



```{r}
options(repr.plot.width = 9, repr.plot.height = 6)
varImpPlot(RF, type = 1)
```

```{r}
rf.pred <- predict(RF, newdata=model_data_2)
```


```{r}
plot(rf.pred, model_data_2$lSalePrice1, main = "Figure 10 Predicted vs. Actual log SalePrice") 
abline(0,1)
```






















































































